Libraries and data uploads
```{r}
library(tidyverse)
library(shiny)
library(shinydashboard)
library(tigris)
library(tidycensus)
library(sf)
library(ggplot2)
library(ggiraph)
library(scales)
library(stringi)
```
```{r}
#Create a data frame with the obesity data
#------------------------------------------
df_n = read_csv("./Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System.csv")%>% 
  select_if(~!all(is.na(.)))
  
#Merge obesity data with map data to create a data frame for interactive map plots
#--------------------------------------------------------------------------------
df_map = merge((get_acs(
  geography = "state",
  variables = "B25077_001",
  year = 2019,
  survey = "acs1",
  geometry = TRUE,
  key = "6640ac063e881a69cd6b2bebd29ab6b6d141923b",
  resolution = "20m"
)%>%shift_geometry()%>%
  dplyr::select(.,-c("estimate","moe"))),df_npa, by.x="NAME",by.y ="LocationDesc" )%>%
  mutate(tooltip = paste(NAME, paste(Data_Value,"%",sep=""), sep = ": "))
```


```{r}
us_value <- get_acs(
  geography = "state",
  variables = "B25077_001",
  year = 2019,
  survey = "acs1",
  geometry = TRUE,
  resolution = "20m"
)%>%shift_geometry()%>%
  dplyr::select(.,-c("estimate","moe"))

write_csv(us_value,"us_map.csv")

us_value = read_csv("./us_map.csv")
```


```{r}
df_npa = read_csv("./Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System.csv")%>% 
  select_if(~!all(is.na(.)))%>%
  rename(demographic = StratificationCategory1, Year = YearEnd, Age = `Age(years)`)%>%
  mutate(demographic = if_else(grepl("Age \\(years\\)" ,demographic), "Age", as.character(demographic)))

df_map = merge(us_value,df_npa, by.x="NAME",by.y ="LocationDesc" )%>%
  mutate(tooltip = paste(NAME, paste(Data_Value,"%",sep=""), sep = ": "))

df_cdi = read_csv("./U.S._Chronic_Disease_Indicators__CDI_.csv")%>% select_if(~!all(is.na(.)))


trend = read_csv("./Obesity vs Diet and Exercise.csv")



```

BASIC EDA OF THE DATA

```{r}
str(df_npa)
head(df_npa)
summary(df_npa)
```


EDA and Graphs
```{r}
df_npa%>%
  filter(LocationDesc=="National",
             Class =="Obesity / Weight Status"
         ,QuestionID =="Q036"
         ,!is.na(Education)
      )%>%
  mutate(Year = as.factor(Year))%>%
  ggplot(aes(y = Data_Value, x = Year))+
  geom_point(aes(color = Education))+
  geom_line()+labs(y = "% Percent of Obesity", 
           title = df_npa$Question[[1]])+
      theme(plot.title = element_text(hjust = 0.5),legend.position="right")#+
  #facet_wrap(~Stratification1)

#unique(df_npa$Class)
```

```{r}
gg = df_npa%>%
  filter(LocationDesc=="National",
         Class =="Obesity / Weight Status",
         QuestionID =="Q036",
         Year==2020,
         !is.na(`Education`)
         )%>%mutate(label_bar=paste(as.character(Data_Value),"%", sep = ""))%>%
  select(`Education`,Data_Value,label_bar,Question)

ggplot(gg ,aes(y = Data_Value, x = `Education`))+
  geom_col(aes(fill =`Education`))+
  geom_text(aes(label = label_bar), vjust = -0.5)+
  labs(y = "% Percent of Obesity", 
       title = gg$Question[[1]])+
  theme(plot.title = element_text(hjust = 0.5),legend.position="none", axis.text.x = element_text(angle = -10))

#ggsave(filename = "file", path = "file/path", width = 8, height = 5)

```





```{r}
(df_cdi%>%
  filter(
    LocationDesc =="United States",
    Topic =="Nutrition, Physical Activity, and Weight Status",
    DataValueType =="Age-adjusted Prevalence",
    StratificationCategory1=="Gender",
    Question=="Obesity among adults aged >= 18 years",
    ))%>%
  arrange(YearEnd,QuestionID)%>%
  
  ggplot(aes(x=factor(YearEnd), y=DataValueAlt))+
  geom_col(aes(fill = Stratification1),position = "dodge")+
  labs(
    title = paste("Percent Obesity Across", df_cdi$StratificationCategory1[[1]]),
    x = "Year", y="Obesity % in Population", 
    fill= df_cdi$StratificationCategory1[[1]])+
  theme(plot.title = element_text(hjust = 0.5),legend.position="none")+
  facet_grid(~Stratification1)


unique((df_cdi%>%
  filter(
    Topic =="Nutrition, Physical Activity, and Weight Status",
    Question == "Median daily frequency of fruit consumption among adults aged >= 18 years"  
  )))

unique(df_cdi$Topic)
#ggsave("Obesity vs Year by Gender.png", width = 8, height = 5 )



```

BASIC EDA OF THE DATA
```{r}
colnames(df_npa)
library(htmlwidgets)

colnames(df_npa)
```


```{r}
# 
# df_npa = read_csv("C:/Users/kellyc/Desktop/NYCDSA/3-R Module/R Shiny Project/Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System.csv")%>% 
#   select_if(~!all(is.na(.)))%>%
#   rename(demographic = StratificationCategory1, Year = YearEnd, Age = `Age(years)`)%>%
#   mutate(demographic = if_else(grepl("Age \\(years\\)" ,demographic), "Age", as.character(demographic)))
# 
# df_npa$Stratification1%>%unique()

# input<-list()
# input$y = 2020
input$d = "Age"

df_npa%>%
      filter(Class =="Obesity / Weight Status",
             QuestionID =="Q036",
             Year==2020,
             demographic == "Age",
             !is.na(demographic)
      )%>%select(Year,LocationDesc,Data_Value,demographic,Stratification1)
  
  
      mutate(label_bar=paste(as.character(Data_Value),"%", sep = ""))%>%
      select(demographic,Data_Value,label_bar,Question,Stratification1)%>%
      ggplot(aes(y = Data_Value, x = Stratification1))+
      geom_col(aes(fill=Stratification1))+
      geom_text(aes(label = label_bar), vjust = -0.5)+
      labs(y = "% Percent of Obesity", 
           title = df_npa$Question[[1]], x=df_npa$demographic[[1]])+
      theme(plot.title = element_text(hjust = 0.5),legend.position="none")

colnames(df_npa)
```

```{r}
df_npa$demographic%>%na.omit()%>%unique()

input$d = "Education"
unique(filter(df_npa,demographic=="Race/Ethnicity",
           QuestionID=="Q036", 
           !is.na(demographic)
           )$Stratification1)

```
```{r}
 df_npa%>%
    filter(
      Year==2020,
      demographic=="Gender",
      Stratification1=="Female",
      QuestionID=="Q036",
      !is.na(demographic))%>%
    select(Year,LocationDesc,Data_Value,demographic,Stratification1)%>%
    rename(States=LocationDesc)
```
```{r}
df_npa%>%count(Total)
```


```{r}
trend$Q047

#na.omit(
  df_npa%>%
  filter(Total == "Total",
         LocationDesc =="National",
         !grepl("overweight",Question))%>%
  pivot_wider(
           names_from = QuestionID,
           values_from = Data_Value)%>%
  arrange(Year)
  #%>%select(Year,Q046))

year = data.frame(Year = c(2011:2020))

for i in colnames()
trend = merge(left_join(year,na.omit(df_npa%>%
  filter(Total == "Total",
         LocationDesc =="National",
         !grepl("overweight",Question))%>%
  pivot_wider(
           names_from = QuestionID,
           values_from = Data_Value)%>%
  arrange(Year)%>%select(Year,Q019)), by = "Year"),trend, on = Year)


write_csv(trend,"Obesity vs Diet and Exercise.csv")
  
```

```{r}
les = df_npa%>%
  filter(Topic=="Physical Activity - Behavior",
         Question =="Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)",
         LocationDesc != "Puerto Rico",
         demographic =="Total")%>%select(Year,LocationDesc,Data_Value)



 inner_join(les,obs, by = c("Year","LocationDesc"))

 ggplot(data = inner_join(les,obs, by = c("Year","LocationDesc")),aes(y=Data_Value.y, x=Data_Value.x))+
  geom_point()+
   geom_smooth(method=lm)+
    labs(title = "Obesity vs Exercise",
         x = "Percent meeting CDC weekly exercise recommendations",
         y="Percent with obesity")

 ggsave("obesity vs Combo.png",width = 8, height = 5)

```


```{r}
les = df_npa%>%
  filter(Topic=="Physical Activity - Behavior",
         Question =="Percent of adults who engage in no leisure-time physical activity",
         LocationDesc != "Puerto Rico",
         demographic =="Total")%>%select(Year,LocationDesc,Data_Value)

obs = df_npa%>%
  filter(Topic=="Obesity / Weight Status",
         Question =="Percent of adults aged 18 years and older who have obesity",
         LocationDesc != "Puerto Rico",
         demographic =="Total")%>%select(Year,LocationDesc,Data_Value)

ggplot(data = inner_join(les,obs, by = c("Year","LocationDesc")),aes(y=Data_Value.y, x=(100-Data_Value.x)))+
  geom_point()+
   geom_smooth(method=lm)+
    labs(title = "Obesity vs General Leisure Time Activity",
         x = "Percent of people active during leisure time.",
         y="Percent of people with obesity")
  
  
ggsave("obesity vs activity.png",width = 8, height = 5)
```


```{r}
les = df_npa%>%
  filter(Question =="Percent of adults who report consuming fruit less than one time daily",
         LocationDesc != "Puerto Rico",
         demographic =="Total")%>%select(Year,LocationDesc,Data_Value)



 inner_join(les,obs, by = c("Year","LocationDesc"))

 ggplot(data = inner_join(les,obs, by = c("Year","LocationDesc")),aes(y=Data_Value.y, x=100-Data_Value.x))+
  geom_point()+
   geom_smooth(method=lm)+
    labs(title = "Obesity vs Fruit Consumption",
         x = "Percent eating fruits daily",
         y="Percent with obesity")

 ggsave("obesity vs fruit.png",width = 8, height = 5)
```


```{r}
les = df_npa%>%
  filter(Question =="Percent of adults who report consuming vegetables less than one time daily",
         LocationDesc != "Puerto Rico",
         demographic =="Total")%>%select(Year,LocationDesc,Data_Value)



 inner_join(les,obs, by = c("Year","LocationDesc"))

 ggplot(data = inner_join(les,obs, by = c("Year","LocationDesc")),aes(y=Data_Value.y, x=100-Data_Value.x))+
  geom_point()+
   geom_smooth(method=lm)+
    labs(title = "Obesity vs Vegetables Consumption",
         x = "Percent eating Vegetables daily",
         y="Percent with obesity")
 
 inner_join(les,obs, by = c("Year","LocationDesc"))$Data_Value.y
 
 
 ggsave("obesity vs veggies.png",width = 8, height = 5)
```


```{r}
```
```{r}

unique(df_npa$Question)

```{r}
df_npa%>%
  filter(Total == "Total",
         LocationDesc =="National",
         !grepl("overweight",Question))%>%
  pivot_wider(
           names_from = QuestionID,
           values_from = Data_Value)%>%
  arrange(Year)


str(trend)

```
```{r}
#Create data frame to graph trends Dietary and Physical Activity Trend vs Obesity

trend_maker = function(df){
  trend = data.frame(cap = c(rep(NA,527)))
  for (cl in unique(df$Class)){
    df1 = df%>%
      filter(LocationDesc!= "National",
             Class == cl,
             demographic =="Total")
    
    for (ques in unique(df1$Question)){
      d=case_when(
        stri_detect_fixed(ques,"obesity")~ "OBS",
        stri_detect_fixed(ques,"300")~ "LongAero",
        stri_detect_fixed(ques,"aerobic physical activity and engage in muscle-strengthening")~ "COMBO",
        stri_detect_fixed(ques,"leisure-time")~ "LESTIME",
        stri_detect_fixed(ques," 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)")~ "ShortAero",
        stri_detect_fixed(ques,"Percent of adults who engage in muscle-strengthening activities on 2 or more days a week")~ "STRENGTH",
        stri_detect_fixed(ques,"fruit")~ "FRUIT",
        stri_detect_fixed(ques,"vegetables")~ "VEGGIE",
        TRUE ~ "CRAP")
      
     df2 = df1%>%
      filter(Question==ques)%>%
      mutate(d = Data_Value)%>%select(Year,d,LocationDesc)
     print(length(df2$Year))
     
     max_length = max(c(length(df2$Year),527))
     
    trend[c(d,paste(d,"Year", sep = "_"),"STATE")] = c(c(df2$d , rep(NA,max_length - length(df2$Year))),c(df2$Year, rep(NA,max_length - length(df2$Year))),c(df2$LocationDesc, rep(NA,max_length - length(df2$Year))))
   
    }
  }
return(trend)

}


trend_maker(df_npa) ->trend


```
```{r}
df_npa%>%
  filter(
    LocationDesc =="National",
    demographic =="Gender",
    QuestionID == "Q036"
    )%>%
ggplot(aes(x = as.character(Year), y = Data_Value ))+
  geom_col(aes(fill = Stratification1),position  = "dodge")+
  facet_grid(~Stratification1)+
  labs(
    title = "Percent Obesity Across Gender",
    x = "Year", y="Obesity % in Population")+
  theme(plot.title = element_text(hjust = 0.5),legend.position="none")


ggsave("Obesity vs Year by Gender.png",width = 8, height = 5)




```

